The Spatial and Temporal Domains of Modern Ecology

### 2016-06-02 paper/main.tex
Full draft as of 1 June 2016

\usepackage{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}

@@

\title{The Spatial and Temporal Domains of Modern Ecology \\
OR Ecology's Spatio-temporal Domains\\
OR Space, Time, and Ecology}

\author
{Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul Elsen$^{4}$, and Tim Treur$^{4}$ \\
\\
\normalsize{$^{1}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{2}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{3}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\
\normalsize{Baltimore, MD 21250, USA}\\
\normalsize{$^{4}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  lestes@princeton.edu.}
}
% Include the date command, but leave its argument blank.

@@

\begin{sciabstract}
An abstract
%  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
\end{sciabstract}

\linenumbers
The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013}.  Ecological patterns emerge within temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}. Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations (e.g. cite) to entire biomes (e.g. cite), while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.

Given the importance of multi-scale studies for providing ecological understanding, and the growing ability to undertake them, it is important to rigorously assess whether ecology is becoming a multi-scale discipline. One approach to answering this question is to quantify the spatial and temporal domains within which observations in ecological studies are collected. Observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992,tilman_ecological_1989}, thus it stands to reason that the temporal and spatial range of ecological observations, and their density within different portions of those ranges, will shed light on modern ecology's progress towards a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}. In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to measure two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration. We collected this information from 367 ecological observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 148 paper subset of 299 randomly selected articles (1.4\% of all papers).

%16 + 29 + 26 + 67 + 14 + 37 + 10
In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority 63\% were collected in plots having resolutions $<$1 m$^2$, while 25\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 12\% in plots of $\geq$1 ha (Fig. 1A). The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 85\% of observations was $<$10 ha, while 31\% covered less than 1 m$^2$ (Fig. 1B).  Only 7\% covered an extent $\geq$10 ha, with just 1.1\% spanning areas $\geq$10 million ha.

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists.pdf}
\vspace{-0.15 cm}
\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies. Bars represent percentage of the 367 collected records falling within each bin.}
\label{afoto1}
\end{figure}

In the temporal dimensions, 30\% of the assessed observations were ``once-offs'' that were not repeated (Fig. 1C). High frequency observations (ranging between as or more frequent than once per second up to daily) comprised 25\% of observations, 20\% were made at daily to monthly time steps, while 35\% and 4\% were respectively made at monthly up to yearly and yearly to decadal intervals.  The temporal duration of studies--the total amount of time the ecological feature was observed (the number of repeat observations multiplied by the effective sampling duration, SI)--was less than 1 day for 59\% of sampled observations,  35\% between 1 day and 1 year, and just 6\% covering greater than 1 year (including several paleoecological studies covering centuries to millenia; Fig. 1D).

The juxtaposition of these observational scales provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (illustrated using two-dimensional kernel density estimates of the log-transformed scale values; Fig. 2). A comparison between the sampling interval and spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig. 1A), the greatest being once-off samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed next by observations of daily to monthly frequency of slightly finer resolution domain (10 cm$^2$ to slightly less than 1 m$^2$), and then a smaller center of monthly to yearly measurements collected at scales of 0.1-10 ha.  Almost no observations were collected with intervals greater than a decade, at all spatial resolutions, while high frequency (daily to sub-daily) observations were rare and collected at very fine resolutions ($<$1000 cm$^2$). Almost no observations with $\geq$1 m$^2$ resolution were collected with sub-second (effectively continuous) to hourly frequency, and the density of 10-1000 m$^2$ resolution, sub-daily to sub-monthly frequency observations was also low.

\begin{figure}[h]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde4.pdf}
\vspace{-0.15 cm}
\caption{}
\label{afoto1}
\end{figure}

In terms of temporal duration and spatial extent (Fig. 1B), the highest frequency ecological observations span time periods of 1 hour to $<$1 month, and cover $\leq$1 m$^2$ of space. The duration-extent domain extent appears to reach its apex at one day of duration and 1000-10,000 ha of extent. The very few observations collected that cover $>$1,000,000 ha were of sub-second duration, as these were collected by satellites that make instantaneous measurements.

Comparing the two spatial scales of measurements against one another shows that observational extent scales log-linearly with resolution (Fig. 1C), revealing a nearly 1:1 correspondence up to 100 ha of both resolution and extent. This relationship indicates that the larger the plot size, the fewer the number of replicates, which is a characteristic of field-based studies because effort and cost typically increases with plot size \cite{kareiva_spatial_1988}. In our sample, 85\% of the observations were collected using manual, field-based methods. These stand in contrast to remotely sensed observations, which made up just 4\% of the sample, but which are characterized by plot resolutions that are very high relative to their spatial extent--these are the low density patches above the 1:1 line defined by the denser concentration of field studies, namely the patches at the following resolution/extent domains; 1-10 m$^2$/100-10,000 ha; 0.1-10 ha/10 million ha; 10-10,000 ha/100,000-10 billion ha.

Comparing sampling interval to temporal duration shows that the bulk of once-off observations last between a minute to a week in duration (Fig. 1D). For repeated observations, the greatest density of observations were those made at monthly to yearly intervals and an aggregate duration of hours to weeks.  Very few repeat observations exceeded a total duration measured in years, and those that did were of daily to yearly frequency.

Reflect nature of observation approaches. \\
Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.  \\
High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
Don't forget to mention automated approaches. 
Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).

Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \

@@

%\begin{scilastnote}
%\item We've included in the template file \texttt{scifile.tex} a new
%environment, \texttt{\{scilastnote\}}, that generates a numbered final
%citation without a corresponding signal in the text.  This environment
%can be used to generate a final numbered reference containing
%acknowledgments, sources of funding, and the like, per {\it Science\/}
%style.
%\end{scilastnote}

@@

>>>>

\usepackage[sort]{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}

@@

\title{The Spatial and Temporal Domains of Modern Ecology }%\\
%OR Ecology's Spatio-temporal Domains\\
%OR Space, Time, and Ecology \\
%OR Ecology: still a small-scale, field-based discipline}

\author
{Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treur$^{4}$ \\
\\
\normalsize{$^{1}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{2}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{3}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\
\normalsize{Baltimore, MD 21250, USA}\\
\normalsize{$^{4}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  lestes@princeton.edu.}
}
% Include the date command, but leave its argument blank.

@@

\begin{sciabstract}
Observational scales plays a critical role in shaping understanding of ecological pattern and process. To gain insight into the spatial and temporal domains being addressed by modern ecological science, we conducted a meta-analysis of papers published in 30 ecological journals between 2004 and 2014. We estimated the spatial resolution and extent and temporal interval and duration of 371 ecological observations reported within 140 papers (out of 364 randomly selected titles). Our analysis revealed that ecology remains a primarily field-based discipline that primarily collects data within narrow spatial and temporal domains. Ecologists still make limited use of methods (e.g. remote sensing, used for 6\% of observations) for observing larger-scale patterns and processes, although make greater use of those which permit high frequency or continuous temporal measurements (12\% of observations). Observational scales were not clearly reported in the majority of three dimensions, which, together with the narrow spatio-temporal ranges of observed scales suggests a limited ability to generalize the findings of many studies in modern ecological research.

%  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
\end{sciabstract}

\linenumbers
The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989-1}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}. Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.

Given the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which ecological research is presently being conducted. To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor). Empirical observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, thus it stands to reason that the temporal and spatial distributions of ecological observations may be indicative of modern ecology's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

We characterized observational domains along two key spatial dimensions, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration, and for each dimension calculated the actual, rather than representative, scales of observation (SI). Our review estimated the spatio-temporal scales of 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (0.85\% of total). In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority (69\%) of observations were collected in plots of $<$1 m$^2$ resolution, 22\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 86\% of observations was $<$10 ha, while 35\% covered less than 1 m$^2$ (Fig. 1B).  Only 14\% covered an extent $\geq$10 ha, with just 4\% spanning areas $\geq$1 million ha.

%16 + 29 + 26 + 67 + 14 + 37 + 10

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists2.pdf}
\vspace{-0.15 cm}
\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies. Bars represent percentage of the 367 collected records falling within each bin.}
\label{afoto1}
\end{figure}

In the temporal dimensions, 36\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 19\% at daily to monthly intervals, 23\% at monthly to yearly intervals, and 5\% at yearly to decadal intervals.  The temporal duration of studies--defined here as the total amount of time the ecological feature was directly observed, and calculated as the number of repeat observations multiplied by the effective sampling duration (SI)--was less than one day for 69\% of sampled observations, between one day and one month for 19\% of observations, between one month and one year for 8\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millenia; Fig. 1D).

The juxtaposition of these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig. 2). Contrasting spatial resolution with sampling interval reveals that the greatest density of observations consists of unreplicated samples collected in plots of 10 cm$^2$ to 1 m$^2$. Two smaller centers of observational concentration are also evident, the largest of which had plots that were also of 10 cm$^2$-1 m$^2$ and were revisited at daily to monthly intervals. A less dense, oblong concentration of observations defined on the lower right by monthly to yearly observations at 100-1000 m$^2$ and on the upper left by near-daily to monthly observations with 0.1-10 ha resolution is also evident. This lower right to upper left orientation marks the tradeoff between resolution and return that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-second intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the inverse tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort and thus reduce sampling frequency \cite{kareiva_spatial_1988}.

%Comparing the spatial and temporal resolutions of  interval to spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig. 2A), the greatest being unreplicated samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed by observations of daily to monthly interval with slightly finer spatial resolutions (10 cm$^2$ to 1 m$^2$), and finally a smaller concentration of monthly to yearly measurements collected at spatial scales of 0.1-10 ha.

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde43.pdf}
\vspace{-0.15 cm}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) sampling interval and spatial resolution, B) temporal duration and spatial extent, C) spatial resolution and spatial extent, and D) sampling interval and temporal duration. Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages. }
\label{afoto1}
\end{figure}

Contrasting temporal duration and spatial extent (Fig. 2B) reveals that the bulk of ecological observations covered total time periods of one hour to $<$1 month and areas of $\leq$1 m$^2$. The next greatest concentration were observations centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent. Very few observations' had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $<$1 hour. In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (i.e. soil cores and weather instrumentation).

Comparing the two spatial dimensions against one another shows that observational extent scales log-linearly with resolution (Fig. 2C). Up to 100 ha, there is a nearly 1:1 correspondence between resolution and extent that results from the tendency for spatial replicates to decline with plot size (and thus effort) in field-based studies \cite{kareiva_spatial_1988}, which accounted for 80\% of the observations in our study. These stand in contrast to most remotely sensed observations, which comprised just 6.5\% of observations, and are characterized by spatial resolutions that are very fine relative to their extent. These constitute the majority of the low density patches (at 0.1-10 ha resolution and 1-10 million ha extent; 100-10,000 ha resolution and 1 million-10 billion ha extent) above the line defined by the denser concentration of field studies in Figure 2C.

A similar linear relationship between sampling interval and temporal duration was not evident. For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to weeks (Fig. 2D). Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals. The majority of unreplicated observations lasted between one minute and one day.

This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI). Observations that were non-continuous in time or space (e.g. point-based field measurements) may effectively represent larger spatial scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}. This concern does not apply to spatially continuous observations, thus the net effect is that if we made estimates of effective spatial scale, the spatial resolution and extent estimates would be compressed from the lower end of their ranges. From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made. Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration. However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}. For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.

Adding to our motivation to record actual observational scales was the fact that most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring. The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases. Although sensitivity analyses show that our general findings are robust to this uncertainty (SI), it would have likely confounded efforts to estimate effective scales.

Despite these caveats, our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research. Our findings suggest that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month). Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, suggesting that, despite well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}. Furthermore, the unclear documenting of observational scales suggests that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}. Taken together, the narrowness of scales and the ambiguity of their documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that much of the knowledge produced in ecology may not be generalizable \cite{margulies_ambiguous_2016,wheatley_factors_2009,wiens_spatial_1989-1}.

The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing. Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this observational method is still under-utilized in ecology.  In contrast, ecologists have more widely adopted methods that increase the temporal resolution of observations. Automated sensors were used to record 12\% of all observations, and accounted for most very high frequency measurements (intervals $\leq$1 hour). As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas it is extremely difficult, if not impossible, to disaggregate long-interval data.

In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains. The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}. Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access. The ever-falling cost of sensor technology and the ubiquity of cell phone connections also means that ecologists have a growing ability to make spatially dense, high frequency observations \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005}. A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.

%Plot resolution issue: need high resolution over large extents, because this relies on remote sensing, and to really capture details, need high resolution.  
%likely because of perceived (and in some cases real) logistical barriers \cite{pettorelli_satellite_2014}.

%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.

%Despite the understanding that ecological patterns and process change with scale \cite{levin_problem_1992}, the narrow distribution of observations within spatial domains calls into question the degree to which ecological phenomena are being investigated at multiple scales. Adding to this concern is the fact the scales of observation were not clearly reported in the majority of studies; the spatial resolution and extent of studies was uncertain in 64\% and 70\% of cases, respectively, while the precise interval and study duration were uncertain in 43\% and 86\% of cases. Although sensitivity analyses show that our conclusions are robust to this uncertainty (SI), the unclear documenting of observational scales suggest that they are not a primary methodological concern in ecological research \cite{chave_problem_2013, wheatley_factors_2009}. 
%
%On the other hand, we found a greater diversity of scales in the temporal domains of observations (Fig. 1D). Over 70\% of sampled observations involved repeated assessments, and one-quarter of these were made with high frequency (intervals $\leq$1 day). Automated, surface-based sensors were used to record most of the very high frequency observations (intervals $\leq$1 hour), and comprised 9\% of all observations, suggesting that ecologists are adopting new technologies for collecting temporal data more rapidly than they are for spatial data.  As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses, whereas it is extremely difficult, if not impossible, to disaggregate long-interval data. 
%
%The temporal duration of most collected observations appears to be brief ($\leq$1year), which is due to the fact that we estimated the summed time of repeated observations, rather than the total time elapsed between the first and last measurements (the span of observation). In many cases, the actual observations are very brief, and thus the temporal duration is narrow. Snapshots in time may be sufficient to capture the temporal dynamics of many phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these our measurement of this domain may understate the temporal breadth of ecological observations. However, for more dynamic phenomena, long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe behavior and evolution of the active fire \cite{clements_observing_2007}. For this reason, and because duration is the temporal analog of spatial extent, we chose to record observational duration rather then span.

%Based on the observational domains recorded here, it appears that ecologists are still primarily collecting data at very fine spatial scales and over very narrow extents. Although the temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics. These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.

%Based on this analysis, ecology thus appears to have

%High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
%Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).

%Although limited sample size, insight into the scales at which science of ecology is making bulk of observations

@@

@@

\begin{scilastnote}
 \item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G),
\end{scilastnote}
